---
title: "Data preparation for AP prediction models"
author: "Bernd Panassiti, Joahnnes Brenner"
date: 'created: 20.07.2015, last modified: `r format(Sys.Date(), format="%d.%m.%Y")`'
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    toc: yes
    toc_depth: 5
  html_document: default
  word_document: default
subtitle: Clustering of monitoring sites
header-includes: \usepackage{graphicx}
---


## Introduction
Due to its topography, for South Tyrol only corse resolution climate date is available that can only be used for sprecies distribution models with caution.
This script creates to two clusters of AP monitoring sites based on their geographical position. Additional clustering criteria are:  
- altitude  
- slope  
- aspect  



```{r setup, warning=FALSE, echo=FALSE}
library(knitr) # for creating html-file
opts_knit$set(root.dir='../') # definining working directory; or normalizePath('../')
opts_chunk$set(fig.align='center') # aligns all figures
opts_chunk$set(echo=FALSE) # suppresses r-code
opts_chunk$set(message=FALSE) # suppresses library outputs
opts_chunk$set(warning=FALSE) # suppresses library warnings
opts_chunk$set(dev='pdf')

```

```{r load}
rm(list=ls(all=TRUE))
source("r-code/00_settings.r")
source("r-code/00_helper_functions.r")
load(file="data/AP_workingdata.Rdata")
set.seed(2015)
```



## Clustering

```{r}
stationsClusterRegion <- stationsCluster[stationsCluster$region %in% sitesCluster$region,]
```



The number of clusters is defined by the number of weather stations, meaning we can rely on k-mean clustering method.

```{r scale}
df <- rbind(sitesCluster,stationsClusterRegion)
df[,3:6] <- scale(df[,3:6])
df <- na.omit(df)
df$region <- as.numeric(df$region)
rownames(df) <- df[,1]
df <- df[,-1]
```


```{r clust_kmeans}
nrClust <- nrow(stationsClusterRegion)
END=FALSE
while(END==FALSE){
clusterFit <- kmeans(df,nrClust) # nrClust = number of weather stations
dfCluster  <- data.frame(df,clusterFit$cluster)
rowsWeatherStations <- which(grepl("st",rownames(dfCluster)))
if(all(seq(1:nrClust) %in% dfCluster[rowsWeatherStations,"clusterFit.cluster"])==TRUE){END=TRUE}
print(nrClust)
nrClust=nrClust-1
}

library(cluster)
library(HSAUR)
dissE <- daisy(df) 
dE2   <- dissE^2
sk2   <- silhouette(clusterFit$cl, dE2)
plot(sk2)
```


```{r}
d <- dist(df, method = "euclidean") # distance matrix
fit <- hclust(d, method="ward") 
plot(fit) # display dendogram
groups <- cutree(fit, k=5) # cut tree into 5 clusters
# draw dendogram with red borders around the 5 clusters 
rect.hclust(fit, k=5, border="red")

library(dendextend)
highlight<-rownames(df)[grepl("st",rownames(df))]
dend <- as.dendrogram(fit)
dend2 <- color_labels(dend, labels=highlight,col=2)
plot(dend2)
groups <- cutree(fit, k=4) # cut tree into 5 clusters
# draw dendogram with red borders around the 5 clusters 
rect.hclust(fit, k=4, border="red")
```

```{r}
clustResult <- data.frame(na.omit(rbind(sitesCluster,stationsClusterRegion))[,1:2],groups)
clustResult[order(clustResult$region),]
```

```{r}
df1 <- clustResult
df2 <- data.frame(sitesCentroid)[c("id","coords.x1","coords.x2")]
df3 <- data.frame(sitesWeatherStations)[,c("NUMMER","coords.x1","coords.x2")]
colnames(df3)[1] <- "id"

dfSites <- merge(df1,df2,by=c("id","id"))
dfSites <- dfSites[order(as.numeric(dfSites[,1])),]
dfSites$type <- "site"
dfStations <- merge(df1,df3,by=c("id","id"))
dfStations$type <- "station"
dfSitesStations <- data.frame(rbind(dfSites,dfStations))

tmp <- SouthTyrol[SouthTyrol@data$name %in% c("Burggrafenamt","Vinschgau"),]
df <- fortify(tmp,region="name")
ggplot(df, aes(long,lat,group=id)) + geom_polygon(fill="white",color="black")+coord_equal() +
  geom_point(data=dfSitesStations,aes(x=coords.x1,y=coords.x2,shape=as.factor(groups),color=type,size=6))


```

