---
title: "Data preparation for AP prediction models"
author: "Bernd Panassiti, Joahnnes Brenner"
date: 'created: 20.07.2015, last modified: `r format(Sys.Date(), format="%d.%m.%Y")`'
output:
  html_document: default
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    toc: yes
    toc_depth: 5
  word_document: default
subtitle: Clustering of monitoring sites
header-includes: \usepackage{graphicx}
---


## Introduction
Due to its topography, for South Tyrol only corse resolution climate date is available that can only be used for sprecies distribution models with caution.
This script creates to two clusters of AP monitoring sites based on their geographical position. Additional clustering criteria are:  
- altitude  
- slope  
- aspect  



```{r setup, warning=FALSE, echo=FALSE}
library(knitr) # for creating html-file
opts_knit$set(root.dir='../') # definining working directory; or normalizePath('../')
opts_chunk$set(fig.align='center') # aligns all figures
opts_chunk$set(echo=FALSE) # suppresses r-code
opts_chunk$set(message=FALSE) # suppresses library outputs
opts_chunk$set(warning=FALSE) # suppresses library warnings
opts_chunk$set(dev='pdf')

```

```{r load}
rm(list=ls(all=TRUE))
source("r-code/00_settings.r")
source("r-code/00_helper_functions.r")
load(file="data/AP_workingdata.Rdata")


if(!require(HSAUR)) 
{
  install.packages("HSAUR", repos = mir)
  require(HSAUR)
}
if(!require(cluster)) 
{
  install.packages("cluster", repos = mir)
  require(cluster)
}
if(!require(ggplot2)) 
{
  install.packages("ggplot2", repos = mir)
  require(ggplot2)
}

if(!require(dendextend)) 
{
  install.packages("dendextend", repos = mir)
  require(dendextend)
}

set.seed(2015)
```



## Clustering

```{r}
stationsClusterRegion <- stationsCluster[stationsCluster$region %in% sitesCluster$region,]
```



```{r scale}
df <- rbind(sitesCluster,stationsClusterRegion)
df[,3:6] <- scale(df[,3:6])
df <- na.omit(df)
df$region <- as.numeric(df$region)
rownames(df) <- df[,1]
df <- df[,-1]
dfClusterRaw <- df # backup
```

### K-means clustering

The number of clusters is defined by the number of weather stations.

```{r clust_kmeans}
df <- dfClusterRaw
nrClust <- nrow(stationsClusterRegion)
END=FALSE
while(END==FALSE){
clusterFit <- kmeans(df,nrClust) # nrClust = number of weather stations
dfCluster  <- data.frame(df,clusterFit$cluster)
rowsWeatherStations <- which(grepl("st",rownames(dfCluster)))
if(all(seq(1:nrClust) %in% dfCluster[rowsWeatherStations,"clusterFit.cluster"])==TRUE){END=TRUE}
#print(nrClust)
nrClust=nrClust-1
}

dissE <- daisy(df) 
dE2   <- dissE^2
sk2   <- silhouette(clusterFit$cl, dE2)
plot(sk2)
```


### Hierarchical clustering

#### Clustering including the region criteria

```{r}
df <- dfClusterRaw
d <- dist(df, method = "euclidean") # distance matrix
fit <- hclust(d, method="ward") 
#plot(fit) # display dendogram
#groups <- cutree(fit, k=5) # cut tree into 5 clusters
# draw dendogram with red borders around the 5 clusters 
#rect.hclust(fit, k=5, border="red")

highlight<-rownames(df)[grepl("st",rownames(df))]
dend <- as.dendrogram(fit)
dend2 <- color_labels(dend, labels=highlight,col=2)
plot(dend2)
groups <- cutree(fit, k=4) # cut tree into 5 clusters
# draw dendogram with red borders around the 5 clusters 
rect.hclust(fit, k=4, border="red")
```

```{r ,results='hide'}
clustResult <- data.frame(na.omit(rbind(sitesCluster,stationsClusterRegion))[,1:2],groups)
clustResult[order(clustResult$region),]
```

```{r}
df1 <- clustResult
df2 <- data.frame(sitesCentroid)[c("id","coords.x1","coords.x2")]
df3 <- data.frame(sitesWeatherStations)[,c("NUMMER","coords.x1","coords.x2")]
colnames(df3)[1] <- "id"

dfSites <- merge(df1,df2,by=c("id","id"))
dfSites <- dfSites[order(as.numeric(dfSites[,1])),]
dfSites$type <- "site"
dfStations <- merge(df1,df3,by=c("id","id"))
dfStations$type <- "station"
dfSitesStations <- data.frame(rbind(dfSites,dfStations))

tmp <- SouthTyrol[SouthTyrol@data$name %in% c("Burggrafenamt","Vinschgau"),]
dfggplot <- fortify(tmp,region="name")
ggplot(dfggplot, aes(long,lat,group=id)) + geom_polygon(fill="white",color="black")+coord_equal() +
  geom_point(data=dfSitesStations,aes(x=coords.x1,y=coords.x2,shape=as.factor(groups),color=type,size=6))
```


Problem: Sites from Vinschgau maybe better assigned to weather stations from Burggrafenamt.



#### Clustering without the region criteria and excluding weather stations from Vinschgau.

```{r}
# without 
df <- dfClusterRaw # backup clustering data.frame
rowSt <- which(grepl("st",rownames(df)) & df[,1]==5)
df <- df[- rowSt,-1]

d <- dist(df, method = "euclidean") # distance matrix
fit <- hclust(d, method="ward") 
```

```{r}
#plot(fit) # display dendogram
#groups <- cutree(fit, k=5) # cut tree into 5 clusters
# draw dendogram with red borders around the 5 clusters 
#rect.hclust(fit, k=5, border="red")

highlight<-rownames(df)[grepl("st",rownames(df))]
dend <- as.dendrogram(fit)
dend2 <- color_labels(dend, labels=highlight,col=2)
plot(dend2)

# draw dendogram with red borders around the 5 clusters 
rect.hclust(fit, k=4, border="red")
```


```{r ,results='hide'}
stationsClusterRegionBackup <- stationsClusterRegion
stationsClusterRegion <- stationsClusterRegion[stationsClusterRegion$region=="Burggrafenamt",]

groups <- cutree(fit, k=4) # cut tree into 5 clusters
clustResult <- data.frame(na.omit(rbind(sitesCluster,stationsClusterRegion))[,1:2],groups)
clustResult[order(clustResult$region),]
stationsClusterRegion <- stationsClusterRegionBackup
```

```{r}
df1 <- clustResult
df2 <- data.frame(sitesCentroid)[c("id","coords.x1","coords.x2")]
df3 <- data.frame(sitesWeatherStations)[,c("NUMMER","coords.x1","coords.x2")]
colnames(df3)[1] <- "id"

dfSites <- merge(df1,df2,by=c("id","id"))
dfSites <- dfSites[order(as.numeric(dfSites[,1])),]
dfSites$type <- "site"
dfStations <- merge(df1,df3,by=c("id","id"))
dfStations$type <- "station"
dfSitesStations <- data.frame(rbind(dfSites,dfStations))

tmp <- SouthTyrol[SouthTyrol@data$name %in% c("Burggrafenamt","Vinschgau"),]
dfggplot <- fortify(tmp,region="name")
ggplot(dfggplot, aes(long,lat,group=id)) + geom_polygon(fill="white",color="black")+coord_equal() +
  geom_point(data=dfSitesStations,aes(x=coords.x1,y=coords.x2,shape=as.factor(groups),color=type,size=6))
```

Problem: Some orchards from Burggrafenamt will not be assigned a weather station when choosing 3 or 4 clusters.  