---
title: "AirTemperature and Precipitation in SouthTyrol | Data Availability and Quality"
author: "Johannes Brenner"
date: "25. Juni 2015"
output: html_document
runtime: shiny
---
  
***
  
This R Markdown document is made interactive using Shiny. To learn more, see [Interactive Documents](http://rmarkdown.rstudio.com/authoring_shiny.html).

***

####Data Origin And Pupose Of Use
 
In South Tyrol a dense network of micro-climate stations is set up since long term. Maintainer of these station in the [Hydrographic Office of the Province Bozen/Bolzano](http://www.provinz.bz.it/wetter/home.asp). Minimum, mean and maximum air temperature and precipitation data has been downloaded on daily basis from their WISKI database. This data is intended to be used for statistical downscaling of regional climate projections, an important step towards realistic forcing for climate impact assessment models.
This document provides an quality check of the whole data set, containing visualisation of gaps in the time series, a visualisation of the time series itself, a suggestion of an interpolated time series, and calculation of quality indices (not included yet).

*** 

```{r, include=FALSE}
if(!require(rgdal)) 
{
  install.packages(rgdal)
  require(rgdal)
}
if(!require(leaflet)) 
{
  install.packages(leaflet)
  require(leaflet)
}
if(!require(zoo)) 
{
  install.packages(zoo)
  require(zoo)
}
if(!require(DataBaseAlpEnvEURAC))
{
  if(!require(devtools))
  {
  install.packages(devtools)
  require(devtools)
  }
  install_github("DataBaseAlpEnvEURAC", "JBrenn")
  require(DataBaseAlpEnvEURAC)
}
if(!require(dygraphs))
{
  install.packages(dygraphs)
  require(dygraphs)
}
if(!require(d3heatmap))
{
  install.packages(d3heatmap)
  require(d3heatmap)
}
```

####Where are the climate stations?

Below one can get an impression of the density of the South Tyrolean meteo network for air temperature. Here available stations providing daily data are shown, high elevation stations with larger radius. An even more dense network could be achieved by adding so far missing stations for which no postprocessing is provided in the Province's data base (e.g. 10min - hourly data).

```{r, echo=FALSE}
# read geographic information
# from file 
# st_Tmax <- read.csv("/run/user/1000/gvfs/smb-share:server=abz02fst.eurac.edu,share=alpenv/Projekte/HiResAlp/06_Workspace/BrJ/02_data/WISKI/Stations_Tmax.csv")
# 
# xy <- project(xy =  cbind(st_Tmax$X_UTM,st_Tmax$Y_UTM), proj = "+proj=utm +zone=32 +ellps=WGS84 +units=m +no_defs", inv = T)
# 
# st_Tmax$long <- xy[,1]; st_Tmax$lat <- xy[,2]
# from DataBseAlpEnv lib

data(st_Tmax)
st_Tmax <- st_Tmax

renderLeaflet({leaflet() %>%
   addProviderTiles("Acetate.terrain") %>%
   addTiles(options = providerTileOptions(opacity = 0.75)) %>%  # Add default OpenStreetMap map tiles
   addCircleMarkers(lng=st_Tmax$long, lat=st_Tmax$lat, popup=paste(st_Tmax$NUMMER, st_Tmax$STATION_D, ", ",  st_Tmax$HOEHE, "m a.s.l"), radius = st_Tmax$HOEHE/100)})
```

***

```{r, echo=FALSE}
# read data from file

# TN_ST <- dB_readZRX2station(files = files, write_csv = T, output_path = "/run/user/1000/gvfs/smb-share:server=abz02fst.eurac.edu,share=alpenv/Projekte/HiResAlp/06_Workspace/BrJ/02_data/WISKI/SouthTyrol/csv/dailyData/", do.hourly = F, do.quality = F, chron = T, multivar = F)

# compute T_mean where missing
# 
# TN_ST_tmean <- lapply(TN_ST, function(x) {
#   if (any(names(x) == "LT_Tmean") | !(any(names(x) == "LT_Tmin") & any(names(x) == "LT_Tmax"))) {
#     return(x)
#   } else {
#       x$LT_Tmean <- (x[,"LT_Tmax"] + x[,"LT_Tmin"]) / 2
#       return(x)
#   }
#   })

# from DataBseAlpEnv lib
data(TN_ST_tmean)
TN_ST_tmean <- TN_ST_tmean
# choose station

inputPanel(
   selectInput(inputId = "station", label = "discover data from station", choices = names(TN_ST_tmean)),
   radioButtons(inputId = "variable", label = "discover", choices = c("Mean daily air temperature","Daily precipitation sums"), selected = "Mean daily air temperature", inline = FALSE)
)

```

***

####Which months do have an adequate data quality?

As for statistical downscaling a time series of at least 20 years is necessary the more general question to ask is: For how many years good quality data is available for the specific station? The heatmap below gives an answer for the choosen station (highlighting of rows and colomns as well as zooming is provided).

```{r, echo=FALSE}
renderD3heatmap({
  if (input$variable=="Mean daily air temperature") {var <- "LT_Tmean"} else {var <- "N_1440"}
  if(input$interpolation=="only raw time series") data <- TN_ST_tmean[[input$station]][,var]
  if(input$interpolation=="add interpolated time series") 
  {
    data <- TN_ST_tmean[[input$station]][,var]
    LT_Tmean_interpol <- na.spline(object = data,  na.rm = F, maxgap = input$maxgap)
    data <-  LT_Tmean_interpol
   }

data_perc <- aggregate(x = data, by = as.yearmon(time(data)), FUN = function(x) {
  y <- sum(!is.na(x)) / length(x) *100
  y <- round(y,0)
})
data_perc <- matrix(c(coredata(data_perc),rep(NA,6)), ncol=12, byrow = T, dimnames = list(1990:2015,month.abb))

d3heatmap(data_perc, scale = "none", colors = "RdYlBu", dendrogram = "none", na.rm = T)
})

#add 
```

***

####How does the series look in detail?

Looking to the series of mean temperature in detail gives you a guess if interpolation of values can help to create sufficent data quality. If you choose _add interpolated time series_ the time series is interpolated with a spline methodology. You can adjust the maximum number of consecutive NAs to fill. Zooming the graph or adding a rolling mean (left bottom corner) to better discover the time series is provided.

```{r, echo=FALSE, warning=FALSE}
inputPanel(
   radioButtons(inputId = "interpolation", label = "show interpolated time series?", choices = c("only raw time series","add interpolated time series"), selected = "only raw time series", inline = FALSE),
   sliderInput(inputId = "maxgap", label = "Consecutive NAs to fill", min = 1, max = 30, value = 4, step = c(1:30))
)

renderDygraph({
  if (input$variable=="Mean daily air temperature") var <- "LT_Tmean" else var <- "N_1440"
  if(input$interpolation=="only raw time series") data <- TN_ST_tmean[[input$station]][,var]
  if(input$interpolation=="add interpolated time series") 
  {
    data <- TN_ST_tmean[[input$station]][,var]
    LT_Tmean_interpol <- na.spline(object = data,  na.rm = F, maxgap = input$maxgap)
    data <- merge(LT_Tmean_interpol,data)
    names(data) <- c("MeanAirTemp_SplineInterpol","MeanAirTemp")
   }
    
  dygraph(data, ylab="degC") %>%
    dyRangeSelector() %>%
    dyRoller()
  
})

```

***