---
title: "Presentation of percentual disease prevalence"
author: "Bernd Panassiti"
date: 'created: 20.02.2015, last modified: `r format(Sys.Date(), format="%d.%m.%Y")`'
output: html_document
header-includes: \usepackage{graphicx}
toc: yes
---


```{r setup, warning=FALSE, echo=FALSE}
library(knitr) # for creating html-file
opts_knit$set(root.dir='../') # definining working directory; or normalizePath('../')
opts_chunk$set(fig.align='center') # aligns all figures
opts_chunk$set(echo=FALSE) # suppresses r-code
opts_chunk$set(message=FALSE) # suppresses library outputs
opts_chunk$set(warning=FALSE) # suppresses library warnings
opts_chunk$set(dev='pdf')

```

```{r load}
rm(list=ls(all=TRUE))
source("r-code/00_settings.r")
source("r-code/00_helper_functions.r")
library(ggplot2)
library(plyr) # function ddply
set.seed(2013)
```

In PostgreSQL:

CREATE TABLE ap_prevalence AS
(Select 
a.id,a.site_id,a.row,a.monitoring_id,a.witchs_broom_2014,  
a.enlarged_stipules_2014,a.small_fruit_2014,a.reddening_2014,a.dense_shoots_2014,   
a.late_blooming_2014,a.new_planting_2014,a.comments_2014,a.cut,a.checkplant,   
a.cultivar,b.region FROM og_appletrees a LEFT JOIN og_parcel_randomorchard b  
ON a.site_id = b.id  
ORDER BY a.id);


Combine Tables og_appletrees

- dense shoots entfernt

INSERT INTO og_appletrees(
site_id,row,monitoring_id,witchs_broom_2014,enlarged_stipules_2014,small_fruit_2014,reddening_2014,
late_blooming_2014,new_planting_2014,comments_2014,cut,checkplant,cultivar,the_geom)
SELECT
site_id,row,monitoring,witchs_bro,enlarged_s,small_frui,reddening_,
late_bloom,new_planti,comments_2,cut,"check",cultivar,the_geom
FROM
og_appletreesdb;



```{r PostgreSQL}
appletreesRegion<-dbGetQuery(con,"select 
site_id,row,id,monitoring_id,witchs_broom_2014,enlarged_stipules_2014,
small_fruit_2014,reddening_2014,dense_shoots_2014,late_blooming_2014,
new_planting_2014,comments_2014,cut,checkplant,cultivar,region
FROM ap_prevalence ORDER BY id")
```



```{r Select_data, include=FALSE}
# get all orchards that are processed -> monitoring_id != NULL
df <- appletreesRegion[!is.na(appletreesRegion$monitoring_id),]

dfVOG <- df[df$region=="Burggrafenamt",]
dfVIP <- df[df$region=="Vinschgau",]
dfVOGGolden <- df[which(df[,16]=="Burggrafenamt" & df[,15] == "Golden"),]
dfVIPGolden <- df[which(df[,16]=="Vinschgau" & df[,15] == "Golden"),]


ApVogVip <- df[which(df[,5]==1 | df[,6]==1 | df[,8]==1),]# all disease rows 
ApVog <- dfVOG[which(dfVOG[,5]==1 | dfVOG[,6]==1 | dfVOG[,8]==1),] # all disease rows Burggrafenamt
ApVip <- dfVIP[which(dfVIP[,5]==1 | dfVIP[,6]==1 | dfVIP[,8]==1),] # all disease rows inschgau

ApVogGolden <- dfVOGGolden[which(dfVOGGolden[,5]==1 | dfVOGGolden[,6]==1 | dfVOGGolden[,8]==1),] # all disease rows Burggrafenamt for Golden Delicious
ApVipGolden <- dfVIPGolden[which(dfVIPGolden[,5]==1 | dfVIPGolden[,6]==1 | dfVIPGolden[,8]==1),] # all disease rows for Vinschgau for Golden Delicious
```

### Disease prevalence

```{r disease_prevalence}
# Select rows with witches broom, enlarged stipulae or reddening
ApVogVipPreval <- nrow(ApVogVip)/nrow(df)*100 # total prevalence (%)
ApVogPreval <- nrow(ApVog)/nrow(dfVOG)*100 # total prevalence (%) for Burggrafenamt
ApVipPreval <- nrow(ApVip)/nrow(dfVIP)*100 # total prevalence (%) for Vinschgau

ApVogGoldenPreval <- nrow(ApVogGolden)/nrow(dfVOGGolden)*100 # total prevalence (%) for Burggrafenamt for Golden Delicious
ApVipGoldenPreval <- nrow(ApVipGolden)/nrow(dfVIPGolden)*100 # total prevalence (%) for Vinschgau for Golden Delicious
```

Disease prevalence for in Burggrafenamt and Vinschgau is `r NumToString(ApVogVipPreval, sigdig=2)`%.  
Disease prevalence for in Burggrafenamt is `r NumToString(ApVogPreval, sigdig=2)`%.  
Disease prevalence for in Vinschgau is `r NumToString(ApVipPreval, sigdig=2)`%.  
Disease prevalence for Golden Delicious in Burggrafenamt is `r NumToString(ApVogGoldenPreval, sigdig=2)`%.  
Disease prevalence for Golden Delicious in Vinschgau is `r NumToString(ApVipGoldenPreval, sigdig=2)`%.  


### Disease prevalence per Site


```{r disease_prevalence_site}
# Select rows with witches broom, enlarged stipulae or reddening
dfSite <-  ddply(df,.(site_id),summarize,length=(length(site_id))) # get all monitored trees for each site

# total prevalence (%)
tmp <- ddply(ApVogVip,.(site_id),summarize,length=(length(site_id)))
ApVogVipSite <- merge(tmp,dfSite,by.x="site_id",by.y="site_id")
ApVogVipSite$per <- (ApVogVipSite[,2]/ApVogVipSite[,3])*100
ApVogVipSiteZero <- nrow(dfSite)-nrow(ApVogVipSite) # Sites with no disease
ApVogVipPer <- mean(c(ApVogVipSite$per,rep(0,ApVogVipSiteZero)))

# total prevalence (%) for Burggrafenamt
tmp <- ddply(ApVog,.(site_id),summarize,length=(length(site_id)))
ApVogSite <- merge(tmp,dfSite,by.x="site_id",by.y="site_id")
ApVogSite$per <- (ApVogSite[,2]/ApVogSite[,3])*100
ApVogSiteZero <- length(unique(dfVOG$site_id))-nrow(ApVogSite)
ApVogPer <- mean(c(ApVogSite$per,rep(0,ApVogSiteZero)))



# total prevalence (%) for Vinschgau
tmp <- ddply(ApVip,.(site_id),summarize,length=(length(site_id)))
ApVipSite <- merge(tmp,dfSite,by.x="site_id",by.y="site_id")
ApVipSite$per <- (ApVipSite[,2]/ApVipSite[,3])*100
ApVipSiteZero <- length(unique(dfVIP$site_id))-nrow(ApVipSite)
ApVipPer <-mean(c(ApVipSite$per,rep(0,ApVipSiteZero)))



# total prevalence (%) for Burggrafenamt for Golden Delicious
tmp <- ddply(ApVogGolden,.(site_id),summarize,length=(length(site_id)))
ApVogGoldenSite <- merge(tmp,dfSite,by.x="site_id",by.y="site_id")
ApVogGoldenSite$per <- (ApVogGoldenSite[,2]/ApVogGoldenSite[,3])*100
ApVogGoldenSiteZero <- length(unique(dfVOGGolden$site_id))-nrow(ApVogGoldenSite)
ApVogGoldenPer <-mean(c(ApVogGoldenSite$per,rep(0,ApVogGoldenSiteZero)))

# total prevalence (%) for Vinschgau for Golden Delicious
tmp <- ddply(ApVipGolden,.(site_id),summarize,length=(length(site_id)))
ApVipGoldenSite <- merge(tmp,dfSite,by.x="site_id",by.y="site_id")
ApVipGoldenSite$per <- (ApVipGoldenSite[,2]/ApVipGoldenSite[,3])*100
ApVipGoldenSiteZero <- length(unique(dfVIPGolden$site_id))-nrow(ApVipGoldenSite)
ApVipGoldenPer <-mean(c(ApVipGoldenSite$per,rep(0,ApVipGoldenSiteZero)))
```

Disease prevalence for in Burggrafenamt and Vinschgau is `r NumToString(ApVogVipPer, sigdig=2)`%.  
Disease prevalence for in Burggrafenamt is `r NumToString(ApVogPer, sigdig=2)`%.  
Disease prevalence for in Vinschgau is `r NumToString(ApVipPer, sigdig=2)`%.  
Disease prevalence for Golden Delicious in Burggrafenamt is `r NumToString(ApVogGoldenPer, sigdig=2)`%.  
Disease prevalence for Golden Delicious in Vinschgau is `r NumToString(ApVipGoldenPer, sigdig=2)`%. 


\pagebreak

### Session info
```{r, echo=FALSE}
devtools::session_info()
```


